ICT and Future Productivity:\\Evidence and Theory of a GPT\thanks{Correspondence: Department of Economics, Boston College, 140 Commonwealth Avenue, Chestnut Hill, MA 02467. Email: brianti@bc.edu (Marco Brianti) and gati@bc.edu (Laura G\'ati).}

### 2018-07-25 Latex/Draft_Summer2018/Summer2018.tex
Progress with writing
We miss (i) interpretation of the results and (ii) conclusions. I personally will wait after the green line for experiments.

\begin{prop}\label{prop:priceRestriction}
In a two-sector economy, if both production functions are identical, then any economy-wide productivity shocks do not affect relative prices between the two sectors.
\end{prop}
\begin{proof}
	See Appendix \ref{section:proofpriceRestriction}.
\end{proof}

@@

\subsubsection{Household}

@@

The production side of the final-good sector consists of a set $[0,1]$ of firms. In each period, final-good firm $j \in [0,1]$ produces $y^c_t(j)$ using the services of labor $l_{t}$, and two types of capital: hard capital, $k^c_{t}$ and ICT capital, $k^i_{t}$. In addition, the production function is augmented with (i) neutral productivity common across sectors, (ii) sector-specific productivity, and (iii) a positive spillover effect related to the diffusion of ICT capital in the whole economy. Production of final output of firm $j$ is undertaken in accordance with
\begin{eqnarray}\label{equation:production_FINAL}
y^c_t(j) = A^c_t \ \big( k^c_{t}(j) \big)^a \ \big( k^i_{t}(j) \big)^b \ \big( l_{t}(j) \big)^{1-a-b}, \ \ 0 < a,b < 1
\end{eqnarray}
where productivity is equal to $ A_t^c = \eta_t \ \theta^c_t \ (k^i_{t})^{\gamma} $. $\eta_t$ is the neutral technological component which moves along the following deterministic trend
\begin{eqnarray}\label{equation:neutral_tech_process}
\eta_t = (\Gamma^{\eta})^t e^{\varepsilon_t},  
\end{eqnarray} 
\begin{eqnarray}\label{equation:neutral_tech_shock}
 \varepsilon_t = \rho_{\nu} \varepsilon_{t-1} + \nu_t
\end{eqnarray} 
where $\rho_{\nu} \in (0,1)$ and $\nu_t \sim N(0,\sigma_{\nu})$ represents a neutral technology shock. $\theta^c_t$ is the sector-specific technological component,  which moves along the following deterministic trend
\begin{eqnarray}
\theta_t^c = (\Gamma^{c})^t.
\end{eqnarray} 
Then, $\gamma$ represents the parameter that governs the size of the spillover effect of ICT capital in the production function. Moreover, $k^c_{t}(j)$ is the part of hard capital $k^c_t$ used by firm $j$ in oder to produce final-good output. Similarly, $k^i_{t}(j)$ and $l_{t}(j)$ are the parts of ICT capital $k^i_t$ and labor $l_t$ used by firm $j$ in oder to produce final-good output, respectively.

@@

Production of ICT capital goods is symmetric to final-output production. It is populated by a set $[0,1]$ of firms. Similarly, ICT-good firm $q \in [0,1]$ produces $y^i_t(q)$ with labor $l_{t}$, hard capital, $k^c_{t}$ and ICT capital, $k^i_{t}$. Analogously to the first sector, production function is augmented by neutral productivity $\eta_t$, sector-specific productivity $\theta^i_t$, and the spillover effect $(k^i_{t})^{\gamma}$. Firm $q$ produces in accordance with
\begin{eqnarray}\label{equation:productionICT}
y^i_t(q) = \eta_t \ \theta^i_t \ (k^i_{t})^{\gamma} \ \ \big( k^c_{t}(q) \big)^a \ \big( k^i_{t}(q) \big)^b \ \big( l_{t}(q) \big)^{1-a-b}, \ \ 0 < a,b < 1
\end{eqnarray}
where $\theta^i_t$ is the sector-specific technological component which moves along the following deterministic trend
\begin{eqnarray}\label{equation:ICT_tech_process}
\theta^i_t = (\Gamma^{i})^t e^{\zeta_t},  
\end{eqnarray} 
\begin{eqnarray}\label{equation:ICT_tech_shock}
\zeta_t = \rho_{\iota} \zeta_{t-1} + \iota_t
\end{eqnarray} 
where $\rho_{\iota} \in (0,1)$ and $\iota_t \sim N(0,\sigma_{\iota})$ represents an ICT technology shock. In line with the first sector, $k^c_{t}(q)$ is the part of hard capital $k^c_t$ used by firm $q$ in oder to produce ICT goods. Similarly, $k^i_{t}(q)$ and $l_{t}(q)$ are the parts of ICT capital $k^i_t$ and labor $l_t$ used by firm $q$ in oder to produce ICT goods, respectively.

@@

We now turn to describing the parameter values and testing the empirical performance of the model. Following \cite{christiano2005nominal}, we partition the model parameters into two groups. We calibrate parameters of the first group using steady-state relations or results from previous works. We estimate the second group by minimizing the distance between the model-implied impulse responses and the empirical results presented in Section \ref{section:mainSpec}. In particular, the second group focuses on parameters specifically related to the effect of ICT investment on future TFP: (i) the size of the spillover effect of ICT capital on productivity $\gamma$, (ii) the standard deviation $\sigma_{\iota}$ and (iii) persistence $\rho_{\iota}$ of ICT shocks.

@@

We calibrate the model at a quarterly frequency. We set the discount factor $\beta$ to $0.97^{\frac{1}{4}}$ to match the average yield of $3$\% of the 3-month Treasury bill using data from 1989:Q1 to 2017:Q1. We set the ICT share $b$ equal to $0.0310$ following \cite{oulton2012long} and hard capital share $a$ equal to $0.3$. Moreover, adapting \cite{oulton2012long} accordingly to our additional assumptions, we derive steady state growth rates of final-sector technology $\Gamma^c$ and ICT-sector technology $\Gamma^i$ from solving the system,
\begin{equation}\label{equation:system_param}
\begin{cases}
\E \big( \frac{p_t}{p_{t-1}} \big) = \frac{\Gamma^c}{\Gamma^i} \\
\E \big( \frac{c_t}{c_{t-1}} \big) = (\Gamma^c)^{\frac{1-b-\gamma}{1-a-b-\gamma}}  (\Gamma^i)^{\frac{b+\gamma}{1-a-b-\gamma}}
\end{cases}
\end{equation}
where $\E[p_t/p_{t-1}]$ and $\E[c_t/c_{t-1}]$ are derived from the simple mean of their empirical related values using quarterly data from 1989:Q1 to 2017:Q1. This implies $\Gamma^c = 1.0034$ and $\Gamma^i = 1.0160$.\footnote{Notice that the value of $\gamma$ is unknown at this step since it is going to be evaluated using impulse-response matching. This implies that the solution for $\Gamma^c$, $\Gamma^i$, and $\gamma$ should be a fixed point where given $\gamma^*$ we obtain $(\Gamma^i)^*$ and $(\Gamma^c)^*$ solving \ref{equation:system_param}. Given $(\Gamma^i)^*$ and $(\Gamma^c)^*$ we obtain $\gamma^*$ by minimizing the impulse-response matching. At this step we are simply solving the system for $\gamma$ equal to zero but a more refine solution will be provided in future.} Finally, to avoid an over-identification problem we set the growth rate of sector-neutral technology $\Gamma^{\eta}$ equal to one.

@@

@@

Prova

@@

>>>>

\begin{prop}\label{prop:priceRestriction}
In a two-sector economy, if both production functions display identical input shares, then any economy-wide productivity shocks do not affect relative prices between the two sectors.
\end{prop}
\begin{proof}
	See Appendix \ref{section:proofpriceRestriction}.
\end{proof}

@@

\subsection{Household}

@@

The production side of the final-good sector consists of a set $[0,1]$ of firms. In each period, final-good firm $j \in [0,1]$ produces $y^c_t(j)$ using the services of labor $l_{t}$, and two types of capital: hard capital, $k^c_{t}$ and ICT capital, $k^i_{t}$. In addition, the production function is augmented with (i) neutral productivity common across sectors, (ii) sector-specific productivity, and (iii) a positive spillover effect related to the diffusion of ICT capital in the whole economy. Production of final output of firm $j$ is undertaken in accordance with
\begin{eqnarray}\label{equation:production_FINAL}
y^c_t(j) = A^c_t \ \big( k^c_{t}(j) \big)^a \ \big( k^i_{t}(j) \big)^b \ \big( l_{t}(j) \big)^{1-a-b}, \ \ 0 < a,b < 1
\end{eqnarray}
where productivity is equal to $ A_t^c = \eta_t \ \theta^c_t \ (k^i_{t})^{\gamma} $. $\eta_t$ is the neutral technological component which moves along the following deterministic trend
\begin{eqnarray}\label{equation:neutral_tech_process}
\eta_t = (\Gamma^{\eta})^t e^{\varepsilon_t},  
\end{eqnarray} 
\begin{eqnarray}\label{equation:neutral_tech_shock}
 \varepsilon_t = \rho_{\nu} \varepsilon_{t-1} + \nu_t
\end{eqnarray} 
where $\rho_{\nu} \in (0,1)$ and $\nu_t \sim N(0,\sigma_{\nu}^2)$ represents a neutral technology shock. $\theta^c_t$ is the sector-specific technological component,  which moves along the following deterministic trend
\begin{eqnarray}
\theta_t^c = (\Gamma^{c})^t.
\end{eqnarray} 
Then, $\gamma$ represents the parameter that governs the size of the spillover effect of ICT capital in the production function. Moreover, $k^c_{t}(j)$ is the part of hard capital $k^c_t$ used by firm $j$ in oder to produce final-good output. Similarly, $k^i_{t}(j)$ and $l_{t}(j)$ are the parts of ICT capital $k^i_t$ and labor $l_t$ used by firm $j$ in oder to produce final-good output, respectively.

@@

Production of ICT capital goods is symmetric to final-output production. It is populated by a set $[0,1]$ of firms. Similarly, ICT-good firm $q \in [0,1]$ produces $y^i_t(q)$ with labor $l_{t}$, hard capital, $k^c_{t}$ and ICT capital, $k^i_{t}$. Analogously to the first sector, production function is augmented by neutral productivity $\eta_t$, sector-specific productivity $\theta^i_t$, and the spillover effect $(k^i_{t})^{\gamma}$. Firm $q$ produces in accordance with
\begin{eqnarray}\label{equation:productionICT}
y^i_t(q) = \eta_t \ \theta^i_t \ (k^i_{t})^{\gamma} \ \ \big( k^c_{t}(q) \big)^a \ \big( k^i_{t}(q) \big)^b \ \big( l_{t}(q) \big)^{1-a-b}, \ \ 0 < a,b < 1
\end{eqnarray}
where $\theta^i_t$ is the sector-specific technological component which moves along the following deterministic trend
\begin{eqnarray}\label{equation:ICT_tech_process}
\theta^i_t = (\Gamma^{i})^t e^{\zeta_t},  
\end{eqnarray} 
\begin{eqnarray}\label{equation:ICT_tech_shock}
\zeta_t = \rho_{\iota} \zeta_{t-1} + \iota_t
\end{eqnarray} 
where $\rho_{\iota} \in (0,1)$ and $\iota_t \sim N(0,\sigma_{\iota}^2)$ represents an ICT technology shock. In line with the first sector, $k^c_{t}(q)$ is the part of hard capital $k^c_t$ used by firm $q$ in oder to produce ICT goods. Similarly, $k^i_{t}(q)$ and $l_{t}(q)$ are the parts of ICT capital $k^i_t$ and labor $l_t$ used by firm $q$ in oder to produce ICT goods, respectively.

@@

We now turn to describing the parameter values and testing the empirical performance of the model. Following \cite{christiano2005nominal}, we partition the model parameters into two groups. We calibrate parameters of the first group using steady-state relations or results from previous works. We estimate the second group by minimizing the distance between the model-implied impulse responses and the empirical results presented in Section \ref{section:mainSpec}. In particular, the second group focuses on parameters specifically related to the effect of ICT investment on future TFP: (i) the standard deviation $\sigma_{\iota}$ of ICT shock $\zeta_t$, (ii) persistence $\rho_{\iota}$ of ICT shock $\zeta_t$ and (iii) the size of the spillover effect of ICT capital on productivity $\gamma$.

@@

We calibrate the model at a quarterly frequency. We set the discount factor $\beta$ to $0.97^{\frac{1}{4}}$ to match the average yield of $3$\% of the 3-month Treasury bill using data from 1989:Q1 to 2017:Q1. We set the ICT share $b$ equal to $0.0310$ following \cite{oulton2012long} and hard capital share $a$ equal to $0.3$. Moreover, adapting \cite{oulton2012long} accordingly to our additional assumptions, we derive steady state growth rates of final-sector technology $\Gamma^c$ and ICT-sector technology $\Gamma^i$ from solving the system,
\begin{equation}\label{equation:system_param}
\begin{cases}
 \frac{p_t}{p_{t-1}}  = \frac{\Gamma^c}{\Gamma^i} \\
 \frac{c_t}{c_{t-1}}  = (\Gamma^c)^{\frac{1-b-\gamma}{1-a-b-\gamma}}  (\Gamma^i)^{\frac{b+\gamma}{1-a-b-\gamma}}
\end{cases}
\end{equation}
where $p_t/p_{t-1}$ and $c_t/c_{t-1}$ are derived from the simple mean of their empirical related values using quarterly data from 1989:Q1 to 2017:Q1. This implies $\Gamma^c = 1.0034$ and $\Gamma^i = 1.0160$.\footnote{Notice that the value of $\gamma$ is unknown at this step since it is going to be evaluated using impulse-response matching. This implies that the solution for $\Gamma^c$, $\Gamma^i$, and $\gamma$ should be a fixed point where given $\gamma^*$ we obtain $(\Gamma^i)^*$ and $(\Gamma^c)^*$ solving \ref{equation:system_param}. Given $(\Gamma^i)^*$ and $(\Gamma^c)^*$ we obtain $\gamma^*$ by minimizing the impulse-response matching. At this step we are simply solving the system for $\gamma$ equal to zero but a more refine solution will be provided in future.} Finally, to avoid an over-identification problem we set the growth rate of sector-neutral technology $\Gamma^{\eta}$ equal to one.

@@

In this section, we estimate model’s key parameters governing the effect of ICT investment on future TFP by imposing theoretical impulse responses of an ICT shock to be as close as possible to the empirical ones. To enhance movement in ICT investment, we use a shock to the ICT-specific technological component $\zeta_t$. Using empirical impulse responses of TFP, ICT investment, consumption, and relative prices, we estimate four parameters: the variance of an ICT technology shock, $\sigma_{\iota}^2$, the persistence of the same shock, $\rho_{\iota}$, and the size of the spillover effect of ICT capital on TFP, $\gamma$. We estimate $\Omega = (\gamma, \ \sigma_{\iota}^2, \ \rho_{\iota})$ by minimizing,
\begin{eqnarray}\label{equation:min_prob_IRmatching}
\min_{\Omega} \big[  \hat{\Psi} - \Psi(\Omega)  \big]' \Lambda \big[  \hat{\Psi} - \Psi(\Omega)  \big]
\end{eqnarray}
where $\Psi(\Omega)$ denotes the mapping from $\Omega$ to the theoretical impulse responses, and $\hat{Psi}$ denotes the empirical impulse responses of an ICT shock to TFP, ICTI, consumption, and relative prices shown in Appendix \ref{section:mainSetResults}. Moreover, $\Lambda$ is a diagonal matrix which embodies variances of the empirical impulse responses. These elements serve as weights to impose relatively more precision on some estimates on the optimization problem. Finally, accordingly to what presented above, we use a horizon of $40$ periods for each response which sum up to a total of $160$ elements for three theoretical parameters only.

Table \ref{table:estimated_parameters} shows results of Problem \ref{equation:min_prob_IRmatching}. The estimate of the spillover effect of ICT capital on TFP, $\gamma$, is $0.58$ indicating that empirical responses favor a causality relation from an ICT capital deepening to total factor productivity. Moreover, estimates of variance and persistence of an ICT-specific technological shock $\sigma^2_{\iota}$ and $\rho_{\iota}$ are $0.01^2$ and $0.9$, respectively. Results imply both a remarkable size and persistence of an ICT sector-specific technological shock $\zeta_t$ in the economy. 
% does it make sense? I dont know what to say for sigmaiota and rhoiota...

Using the estimated parameter values in Table \ref{table:estimated_parameters}, Figure \ref{fig:IR_matching} displays empirical impulse responses along with theoretical ones. Our simple model is able to mimic empirical movements in TFP, ICT investment, consumption, and relative prices. In both the model and the data, a sector-specific technological shock $\zeta_t$ has a delayed positive effect on TFP which tend to last over the medium and long run. Moreover, in both cases, ICT investment has a large initial peak and then tends to remain persistently positive over the $40$-quarter horizon. In the model and the data, consumption responds positively on impact, reaches its peak after $4$ years and persistently remains positive. Finally, model-generated and empirical impulse responses of relative prices have a large and negative impact response which dies out in about $6$ years.

@@

Notice that in this proof we use the same assumptions and the same notation in the 2-sector model presented in Section \ref{section:theory}.

First order condition with respect to labor input for firm $j$ is
\begin{eqnarray}\label{equation:FOC_j_l}
w_t =  (1 - a - b) p_t \eta_t \theta^c (k^i_t)^{\gamma} \bigg( \frac{k_t^c(j)}{l_t(j)} \bigg)^a \bigg( \frac{k_t^i(j)}{l_t(j)} \bigg)^b,
\end{eqnarray}
and first order condition with respect to labor input for firm $q$ is
\begin{eqnarray}\label{equation:FOC_q_l}
w_t = (1 - a - b) \eta_t \theta^c (k^i_t)^{\gamma} \bigg( \frac{k_t^c(q)}{l_t(q)} \bigg)^a \bigg( \frac{k_t^i(q)}{l_t(q)} \bigg)^b,
\end{eqnarray}
Since both production functions display identical input shares then in equilibrium input ratios must be the same over the two sectors, which is,
\begin{eqnarray}\label{eq:shares_result1}
\frac{k_t^c(j)}{l_t(j)} = \frac{k_t^c(q)}{l_t(q)},
\end{eqnarray}
\begin{eqnarray}\label{eq:shares_result2}
\frac{k_t^i(j)}{l_t(j)} = \frac{k_t^i(q)}{l_t(q)},
\end{eqnarray}
\begin{eqnarray*}\label{eq:shares_result3}
\frac{k_t^i(j)}{k^c_t(j)} = \frac{k_t^i(q)}{k^c_t(q)}.
\end{eqnarray*}
Now, dividing \ref{equation:FOC_j_l} over \ref{equation:FOC_q_l}, we can use \ref{eq:shares_result1} and \ref{eq:shares_result2} to obtain
\begin{eqnarray}\label{eq:result_proof}
1 = \frac{1}{p_t} \frac{\theta^c_t}{\theta^i_t} \ \ \Rightarrow \ \ p_t = \frac{\theta^c_t}{\theta^i_t}.
\end{eqnarray}
From Equation \ref{eq:result_proof}, it is clear that the equilibrium path of $p_t$ is fully independent on the past, current, and future values of neutral technology, $\eta_t$. 
\qed

@@

\section{Estimated Parameters via IR Matching}

 	\begin{table}[h!]
	\begin{center}
		\begin{tabular}{clc}
			\hline
			Symbol & Economic Interpretation & Estimated Value \\
			\hline
            $\sigma^2_{\iota}$ & Variance of ICT technological shock &   $0.01^2$  \\
            $\rho_{\iota}$     & Persistence of ICT technological shock & $0.9$   \\
            $\gamma$           & Size of spillover of ICT capital on TFP &  $0.5881$ \\
			\hline
		\end{tabular}
		\caption{Estimation of key parameters by imposing model's impulse responses of an ICT shock to be as close as possible to the empirical ones. Theoretical estimates are impulse responses to an ICT-specific technological shock, $\zeta_t$ which are functions of key parameters: $\sigma_{\iota}^2$, $\rho_{\iota}$, and $\gamma$. Empirical impulse responses of an ICT shock to TFP, ICTI, consumption, and relative prices are derived from Specification \ref{eq:mainSpecification} using the empirical strategy presented in Section \ref{section:empiricalstrategy_simple}. According to what presented in Appendix \ref{section:mainSetResults}, we use a horizon of $40$ periods for each response.}
		\label{table:estimated_parameters}
	\end{center}
\end{table}

\section{Impulse-Response Matching}\label{section:appendix_IR_matching}

\begin{figure}[h!]
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{MainFigures/fig_ICT_Shock_on_TFP_IRmatching_together}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{MainFigures/fig_ICT_Shock_on_ICTI_IRmatching_together}
	\end{subfigure}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{MainFigures/fig_ICT_Shock_on_C_IRmatching_together}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{MainFigures/fig_ICT_Shock_on_RP_IRmatching_together}
	\end{subfigure}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\caption{Empirical VAR versus model responses to an ICT shock. The black solid lines show the empirical responses to an identified ICT shock in Specification \ref{eq:mainSpecification}. The dotted red lines show the model’s responses to a shock to the ICT-specific technological component $\zeta_t$. The horizontal axes refer to forecast horizon and the units of the vertical axes are percentage deviations.}
	\label{fig:IR_matching}
\end{figure}
